Cycles of the logistic map

### 2013-05-14 doc/pc-ijbc.tex
pc-ijbc: improve sec. 2.7

\newcommand{\vx}{\vct x}
\newcommand{\vxn}[1]{\vx^{(#1)}}
\newcommand{\Tr}{\mathrm{Tr}}
% basis set of cyclic polynomials
\newcommand{\Pset}{\mathcal P}
\newcommand{\NB}{N_\Pset}
\newcommand{\LB}{L_\Pset}
\newcommand{\Rlam}{(R, \lambda)}

@@

Cyclic polynomials
  generated from monomials of unit coefficient
  (such as $a$ and $b$, but not $a + b$ in the above example)
  can be labeled as follows.
We write the monomial generator as
  ${x_1}^{e_1} {x_2}^{e_2} \dots {x_n}^{e_n}$,
  and form a sequence $p$ of indices with
  $e_1$ 1's, $e_2$ 2's, \ldots, $e_n$ $n$'s.
The sequence $p$ serves as the label of the cyclic polynomial $C_p(\vxn{n})$,
  where $\vxn{n} \equiv \{x_1, \ldots, x_n\}$
[we will, however, drop the superscript ``$^{(n)}$'' or
  even the argument ``$(\vxn{n})$'' entirely when it is clear].
For example,
  $C_{12}(\vx)  = x_1 x_2 + x_2 x_3 + \dots + x_n x_1$ and
  $C_{112}(\vx) = {x_1}^2 x_2 + {x_2}^2 x_3 + \dots + {x_n}^2 x_1$.
In case the cyclic polynomial has several generators,
  we pick the one with the smallest $p$
  in the sense of lexicographic order,
  e.g., we choose $x_1 x_2$ instead of $x_2 x_3$ for $C_{12}(\vx)$
  because $12 < 23$ ($n \ge 3$).
Finally, we add $C_0(\vx) \equiv 1$ for completeness.
%
%
%

@@

\subsection{\label{sec:primfac}Minimal polynomial of the $n$-cycles}

@@

%\begin{theorem}
%  The polynomial $P_n\Rlam$ of all $n$-cycles
%  is a factor of the characteristic polynomial $A_n\Rlam$,
%  and can be computed as
  \begin{equation}
    P_n\Rlam
    = \prod_{cd = n} B_d^c\Rlam^{\mu(c)},
  \label{eq:primfac}
  \end{equation}
where
  $B_d^c\Rlam \equiv \prod_{k=1}^c A_{d}(R, e^{2k\pi i/c} \lambda^{1/c})$,
  $\lambda^{1/c}$ is a complex $c$th root of $\lambda$.
%\end{theorem}
%
%
%\begin{remark}
%Here, ``minimal'' means that the polynomial contains all zeros.
%It happens to be irreducible for a general $\lambda$.
%At $\lambda = 1$, the polynomial of $R$, however, can be factorized (see later).
%\end{remark}
%
%\begin{remark}
The M\"obius function $\mu(n)$ is $(-1)^k$
  if $n$ is the product of $k$ distinct primes,
  or 0 if $n$ is divisible by a square of a prime.
%Thus,
The first few $\mu(n) = 1, -1, -1, 0, -1, 1, -1, \ldots,$
  from $n = 1$%.
%The $\mu(n)$ is useful for inversion:
%$g(n) = \sum_{d|n} \mu(n/d) \, h(d)$
%if and only if $h(n) = \sum_{d|n} g(d)$
\cite{hardy}.
%\end{remark}
%
%
\begin{remark}
%$B_{d,c}\Rlam$ is a polynomial of $\lambda$.
Despite the argument $\lambda^{1/c}$,
  the product $\prod_{k=1}^c A_d(R, e^{2k\pi i/c} \lambda^{1/c})$,
  hence $B_d^c\Rlam$,
  is free from radicals of powers of $\lambda^{1/c}$,
  for it is invariant under $\lambda \rightarrow e^{2\pi i} \lambda$.
Thus, $B_d^c\Rlam$ is a polynomial of $\lambda$,
and $\deg_\lambda B_d^c\Rlam = \deg_\lambda A_d\Rlam$.
Particularly, $B_n^1\Rlam = A_n\Rlam$.
\end{remark}
%
%
%We will refer to $P_n\Rlam$ as the \emph{primitive factor} below.
%
%
Let us apply \refeq{primfac}
to the examples in \refsec{examples}.
We will omit the argument $R$ below for convenience.
%%
For $n = 1$,
%there is no irrelevant factor in \refeq{xr1} and
$P_1(\lambda) = B_1^1(\lambda) = A_1(\lambda)
  = \lambda^2 - 2\,\lambda - 4\,R$.
%
%
%
For $n = 2$, since
$B_1^2(\lambda)
=
A_1(\sqrt{\lambda})
A_1(-\sqrt{\lambda})
=
(4R-\lambda)^2 -4\lambda$,
%
$P_2(\lambda) = A_2(\lambda) \, B_1^2(\lambda)^{-1} = 4R - 4 + \lambda$.
%
%
For $n = 4$,
%we have
%$16 (R-1)^2 - \lambda
%= (4R - 4 + \sqrt{\lambda})(4R - 4 - \sqrt{\lambda})$
%and
%$
%(16 R^2 + \lambda)^2 - 16 (2R+1)^2 \lambda
%=
%\big[(4R - \sqrt{\lambda})^2  - 4\sqrt{\lambda}\,\big]
%\big[(4R + \sqrt{\lambda})^2  + 4\sqrt{\lambda}\,\big]$.
%Thus,
we can show that
the last two factors of \refeq{xr4}
can be written as
$B_2^2(\lambda) = A_2(\sqrt{\lambda}) A_2(-\sqrt{\lambda})$,
and
%
$P_4(\lambda)
    = A_4(\lambda) \, B_2^2(\lambda)^{-1}$
indeed the yields the sextic factor.
%$    = 4096 R^6-12288 R^5+ 256 (\lambda + 48) (R^4- R^3)
%    -16(\lambda + 32)(\lambda - 16) R^2 - (\lambda-16)^3$.
Note, $B_1^4(\lambda)$ is unused for $\mu(4) = 0$.

@@

%We will prove \refthm{primfac} by the following steps.
%
To show \refeq{primfac}, we first find the degrees in $\lambda$ of
  $A_n\Rlam$ [\refeq{necklace}] and $P_n\Rlam$ [\refeq{lyndon}].
By comparing the degrees,
  we then show that each $P_d\Rlam$ with ($d|n$),
  after some transformation,
  contributes one polynomial factor to $A_n\Rlam$
  [\refeq{prod}],
  and the inversion of the relation
  yields \refeq{primfac}.

@@

First, to count the square-free cyclic polynomials,
  we notice a one-to-one mapping
  from the square-free cyclic polynomials
  to the binary necklaces.
%
A binary necklace
  is a nonequivalent binary $0$-$1$ string.
%
Two strings are equivalent if they differ only by a circular shift.
%
For example, for $n = 3$, there are $2^3 = 8$ binary strings,
but only four necklaces: 000, 001, 011 and 111,
since 010 and 100 are equivalent to 001,
so are 110 and 101 to 011.
%
The mapping is constructed as follows.
%
Given a cyclic polynomial, if the generator contains $x_k$,
  the $k$th character from the left of the necklace is 1,
  otherwise 0,
e.g. the above four necklaces correspond to
  $C_0$, $C_1$, $C_{12}$ and $C_{123}$, respectively.
%For example, if $n = 3$,
% the square-free polynomials for $000$, $100$, $110$ and $111$ are
%  $C_0 = 1$ (generator: 1),
%  $C_1 = x_1 + x_2 + x_3$ (generator: $x_1$),
%  $C_{12} = x_1 x_2 + x_2 x_3 + x_3 x_1$ (generator: $x_1 x_2$)
%  and $C_{123} = x_1 x_2 x_3$ (generator: $x_1 x_2 x_3$), respectively.
Thus,
%
%\begin{theorem}
the number $\NB(n)$ of the square-free cyclic polynomials $C_p$,
  %$(\vx)$ formed by the cycle points $\vx = \{x_1, \ldots, x_n\}$,
  is given by
\begin{equation}
  %\deg_\lambda A_n\Rlam =
  \NB(n) = (1/n) \textstyle\sum_{d|n} \varphi(n/d) 2^{\,d}.
\label{eq:necklace}
\end{equation}
%
%\label{thm:necklace}
%\end{theorem}
%
Here, $\varphi(m)$ is Euler's totient function,
which gives the number of integers from 1 to $m$
  that are coprime to $m$,
  e.g.,
  $\varphi(1) = 1$,
  $\varphi(3) = 2$ for 1 and 2,
  $\varphi(6) = 2$ for 1 and 5.
By \refeq{xr},
  $\NB$ is also $\deg_\lambda A_n\Rlam$.
The first few $\NB(n)$ are
2, 3, 4, 6, 8, 14, 20, 36, 60, 108, \dots, % 188, 352, 632, \ldots,
  from $n = 1$.

@@

%\begin{theorem}
Next, the degree in $\lambda$ of the polynomial $P_n\Rlam$ of all $n$-cycles
  is equal to the number of the $n$-cycles,
  which is given by
\begin{equation}
  %\deg_\lambda P_n\Rlam =
  \LB(n) = (1/n)\textstyle\sum_{d | n} \mu(n/d) 2^{\,d}.
\label{eq:lyndon}
\end{equation}
%\label{thm:lyndon}
%\end{theorem}
%
%
%\begin{proof}
The number $\LB$ of $n$-cycles
  is well known \cite{hao, hao2, lutzky}.
The minimal polynomial $P_n\Rlam$
  can be written as $\prod_c (\lambda - \lambda_c)$,
where $\lambda_c$ is the value of $\Lambda(\vx^c)$
  evaluated at the $c$th cycle.
Thus, $\deg_\lambda P_n\Rlam = \LB$.
%\end{proof}
%
The first few $\LB$ are
2, 1, 2, 3, 6, 9, 18, 30, 56, 99, \dots %186, 335, 630, \ldots,
from $n = 1$.

Incidentally, $\LB$ is also the number of aperiodic necklaces of length $n$
(e.g., for $n = 4$,
  $0011$ is aperiodic, but $0101$ is not due to the shorter period 2).
%
Since an $n$-necklace is either aperiodic itself or
  a repeated aperiodic $d$-necklace (with $d|n$),
  we have
\begin{equation}
  \NB(n) = \textstyle\sum_{d|n} \LB(d).
\label{eq:necklacelyndon}
\end{equation}
%
Since the left side is also $\deg_\lambda A_n\Rlam$,
  and $A_n\Rlam$ has to represent all $d$-cycles with $d|n$,
$A_n\Rlam$
%and each cycle holds a distinct $\lambda$,
%  $\deg_\lambda A_n\Rlam$
%  is at least $\sum_{d|n} L(d)$
%  according to \refeq{lyndon},
%  which is equal to
%  $\NB$ by \refeq{necklacelyndon}.
%Since $\deg_\lambda A_n\Rlam = \NB$ by \refeq{necklace},
%  each $n$-cycle occurs exactly once in $A_n\Rlam$.

%\subsubsection{Relation between $P_n$ and $A_n$}

%\begin{theorem}
We can now show that $P_n\Rlam$ and 
and $A_n\Rlam$ are related by
  \begin{equation}
    A_n\Rlam = \prod_{cd = n} Q_d^c\Rlam,
    \label{eq:prod}
  \end{equation}
  where
  $Q_d^c\Rlam = \prod_{k=1}^{c} P_d(R, e^{2 k \pi i/c} \lambda^{1/c})$
  is a polynomial of degree $L(d)$ in $\lambda$
  representing contributions from $d$-cycles.
%encompassed by $A_n\Rlam$.
%\label{thm:prod}
%\end{theorem}

%\begin{proof}
In a $d$-cycle,
 we have $\Lambda(\{x_1, \ldots, x_n\}) = (-2)^n x_1 \dots x_n
   = {\Lambda_d}^c$,
where
$\Lambda_d(\{x_1, \ldots, x_d\}) \equiv (-2)^d x_1 \dots x_d$
and
$c \equiv n/d$.
So the $d$-cycle satisfies a polynomial
  $P_d(R, \lambda^{1/c}) = 0$,
where $\lambda = \Lambda(\{x_1,\ldots,x_n\})$.
%
This is, however, not a polynomial equation, and
%
the radical $\lambda^{1/c}$ can be removed by the product
$Q_d^c\Rlam
  \equiv \prod_{k=1}^c P_d(R, e^{2k\pi i/c} \, \lambda^{1/c}) = 0$.
%
Now $Q_d^c\Rlam$ is a polynomial of $\lambda$
  for it is invariant under
  $\lambda \rightarrow e^{2\pi i} \lambda$,
and thus free from radicals of the form $\lambda^{l/c}$
  \big[if $(l, c) \ne c$\big].
And since
  $\deg_\lambda Q_d^c\Rlam
    = \deg_\lambda P_d\Rlam = L(d)$,
  it is also a polynomial
  of the lowest possible degree in $\lambda$.

Therefore, the product $\prod_{cd=n} Q_d^c\Rlam$
  can differ from $A_n\Rlam$ only by
  a multiple.
Since $Q_n^1\Rlam = P_n\Rlam$,
  and the coefficient of highest power of $\lambda$ is always
  unity in $A_n\Rlam$ \big[see the definition \refeq{xr}\big],
  we know by induction that the coefficients of the highest power of $\lambda$
  in all $P_n\Rlam$ and $Q_d^c\Rlam$
  are also unities.
So the multiple is one, hence \refeq{prod}.
%
%
%
%\end{proof}

\refeq{primfac} is a corollary of \refeq{prod},
\begin{small}
\begin{align*}
 B_n^m (\lambda)
% & = \prod_{l = 1}^{m} A_n(e^{2l\pi i/m} \lambda^{1/m}) \\
 & = \prod_{c | n} \prod_{l=1}^m \prod_{k=1}^c
    P_{n/c}(e^{\frac{2k\pi i}{c} + \frac{2l\pi i}{m c}} \lambda^{1/(m c)}) \\
 & = \prod_{c | n} \prod_{k'=1}^{m c}
    P_{n/c}(e^{2k'\pi i/(m c)} \lambda^{1/(m c)})% \\
 %&
  = \prod_{c | n} Q_{n/c}^{m c}(\lambda).
\end{align*}
\end{small}
The M\"obious inversion yields
\begin{equation}
Q_n^m\Rlam
= \prod_{d|n} B_d^{m n/d}\Rlam^{\mu(n/d)},
\notag
%\tag{$\ref{eq:primfac}'$}
\end{equation}
which is reduced to \refeq{primfac} with $m = 1$,
for $Q_n^1 = P_n$.

@@

Since the $n$-odd-cycles satisfy a polynomial of lower degree,
  the minimal polynomial $P_{2n}(r, \lambda)$
  can always be factorized.
%
Suppose $\Lambda^\odd(\vx) \equiv \prod_{k = 1}^{n} f'(x_k)$ in the odd-cycle
 satisfies $P_{n}^\odd(r, \lambda^\odd) = 0$
\big[where $\lambda^\odd$ is the value of $\Lambda^\odd(\vx)$,
and $\lambda^\odd = \pm\sqrt \lambda$\,\big],
then $P^\odd_{n}(r, \sqrt \lambda) P^\odd_{n/2}(r, -\sqrt \lambda)$
is a factor of $P_{n}(r, \lambda)$.
%
For example, by solving the $n = 2$ odd-cycle,
  we have $- x_1 = r x_1 - {x_1}^3$, or ${x_1}^2 = r + 1$.
Since $\lambda^\odd = r - 3 {x_1}^2$,
$P^\odd_1(r, \lambda^\odd) = \lambda^\odd + 2 r + 3$.
%
The minimal polynomial of 2-cycles
  $P_2(r,\lambda) = -\big[(2r+3)^2 - \lambda\big]\,(\lambda + 2r^2-9)$,
  indeed contains
  $P^\odd_1(r, \sqrt \lambda) \, P^\odd_1(r, -\sqrt \lambda)
   = (\sqrt \lambda + 2 r + 3)(-\sqrt \lambda + 2 r + 3)$
  as a factor.

>>>>

\newcommand{\vx}{\vct x}
\newcommand{\vxn}[1]{\vx_{#1}}
\newcommand{\Tr}{\mathrm{Tr}}
% basis set of cyclic polynomials
\newcommand{\Pset}{\mathcal P}
\newcommand{\NB}{N_\Pset}
\newcommand{\LB}{L_\Pset}
\newcommand{\lam}{(\lambda)}
\newcommand{\Rlam}{(R, \lambda)}

@@

Cyclic polynomials
  generated from monomials of unit coefficient
  (such as $a$ and $b$, but not $a + b$ in the above example)
  can be labeled as follows.
We write the monomial generator as
  ${x_1}^{e_1} {x_2}^{e_2} \dots {x_n}^{e_n}$,
  and form a sequence $p$ of indices with
  $e_1$ 1's, $e_2$ 2's, \ldots, $e_n$ $n$'s.
The sequence $p$ serves as the label of the cyclic polynomial $C_p(\vxn{n})$,
  where $\vxn{n} \equiv \{x_1, \ldots, x_n\}$
[we will, however, drop the subscript ``$_n$'' or
  even the argument ``$(\vxn{n})$'' entirely when it is clear].
For example,
  $C_{12}(\vx)  = x_1 x_2 + x_2 x_3 + \dots + x_n x_1$ and
  $C_{112}(\vx) = {x_1}^2 x_2 + {x_2}^2 x_3 + \dots + {x_n}^2 x_1$.
In case the cyclic polynomial has several generators,
  we pick the one with the smallest $p$
  in the sense of lexicographic order,
  e.g., we choose $x_1 x_2$ instead of $x_2 x_3$ for $C_{12}(\vx)$
  because $12 < 23$ ($n \ge 3$).
Finally, we add $C_0(\vx) \equiv 1$ for completeness.
%
%
%

@@

\subsection{\label{sec:primfac}$n$-cycle polynomial}

@@

%\begin{theorem}
%  The polynomial $P_n\Rlam$ of all $n$-cycles
%  is a factor of the characteristic polynomial $A_n\Rlam$,
%  and can be computed as
  \begin{equation}
    P_n\Rlam
    = \prod_{cd = n} B_d^c\Rlam^{\mu(c)},
  \label{eq:primfac}
  \end{equation}
where
  $B_d^c\Rlam \equiv \prod_{k=1}^c A_{d}(R, e^{2k\pi i/c} \lambda^{1/c})$,
  $\lambda^{1/c}$ is a complex $c$th root of $\lambda$.
%\end{theorem}
%
%
%\begin{remark}
%Here, ``minimal'' means that the polynomial contains all zeros.
%It happens to be irreducible for a general $\lambda$.
%At $\lambda = 1$, the polynomial of $R$, however, can be factorized (see later).
%\end{remark}
%
%\begin{remark}
The M\"obius function $\mu(n)$ is $(-1)^k$
  if $n$ is the product of $k$ distinct primes,
  or 0 if $n$ is divisible by a square of a prime.
%Thus,
The first few $\mu(n) = 1, -1, -1, 0, -1, 1, -1, \ldots,$
  from $n = 1$
%The $\mu(n)$ is useful for inversion:
%$g(n) = \sum_{d|n} \mu(n/d) \, h(d)$
%if and only if $h(n) = \sum_{d|n} g(d)$
 \cite{hardy}.
%\end{remark}
%

Since the discussions in this and the next sections
  concern $\lambda$ instead of $R$,
  we will drop the argument $R$ in $P_n\Rlam$ and $A_n\Rlam$, etc.
%\begin{remark}
%$B_{d,c}\lambda$ is a polynomial of $\lambda$.
Despite the argument $\lambda^{1/c}$,
  the product $\prod_{k=1}^c A_d(e^{2k\pi i/c} \lambda^{1/c})$,
  hence $B_d^c\lam$,
  is free from radicals of powers of $\lambda^{1/c}$,
  for it is invariant under $\lambda \rightarrow e^{2\pi i} \lambda$.
Thus, $B_d^c\lam$ is a polynomial of $\lambda$,
and $\deg_\lambda B_d^c\lam = \deg_\lambda A_d\lam$.
Particularly, $B_n^1\lam = A_n\lam$.
%
%\end{remark}

%
Let us apply \refeq{primfac}
to the examples in \refsec{examples}.
%%
For $n = 1$,
%there is no irrelevant factor in \refeq{xr1} and
$P_1(\lambda) = B_1^1(\lambda) = A_1(\lambda)
  = \lambda^2 - 2\,\lambda - 4\,R$.
%
%
%
For $n = 2$, since
$B_1^2(\lambda)
=
A_1(\sqrt{\lambda})
A_1(-\sqrt{\lambda})
=
(4R-\lambda)^2 -4\lambda$,
%
$P_2(\lambda) = A_2(\lambda) \, B_1^2(\lambda)^{-1} = 4R - 4 + \lambda$.
%
%
For $n = 4$,
%we have
%$16 (R-1)^2 - \lambda
%= (4R - 4 + \sqrt{\lambda})(4R - 4 - \sqrt{\lambda})$
%and
%$
%(16 R^2 + \lambda)^2 - 16 (2R+1)^2 \lambda
%=
%\big[(4R - \sqrt{\lambda})^2  - 4\sqrt{\lambda}\,\big]
%\big[(4R + \sqrt{\lambda})^2  + 4\sqrt{\lambda}\,\big]$.
%Thus,
we can show that
the last two factors of \refeq{xr4}
can be written as
$B_2^2(\lambda) = A_2(\sqrt{\lambda}) A_2(-\sqrt{\lambda})$,
and
%
$P_4(\lambda)
    = A_4(\lambda) \, B_2^2(\lambda)^{-1}$
indeed the yields the sextic factor.
%$    = 4096 R^6-12288 R^5+ 256 (\lambda + 48) (R^4- R^3)
%    -16(\lambda + 32)(\lambda - 16) R^2 - (\lambda-16)^3$.
Note, $B_1^4(\lambda)$ is unused for $\mu(4) = 0$.

@@

%We will prove \refthm{primfac} by the following steps.
%
We now prove \refeq{primfac}.
%
To do so, we first find the degrees in $\lambda$ of
  $A_n\lam$ [\refeq{necklace}] and $P_n\lam$ [\refeq{lyndon}].
By comparing the degrees,
  we then show that each $P_d\lam$ with $d|n$,
  after some transformation,
  contributes one polynomial factor to $A_n\lam$
  [\refeq{prod}],
  and the inversion of the relation
  yields \refeq{primfac}.

@@

First, to count the square-free cyclic polynomials,
  we notice a one-to-one mapping
  from the square-free cyclic polynomials
  to the binary necklaces.
%
A binary necklace
  is a nonequivalent binary $0$-$1$ string.
%
Two strings are equivalent if they differ only by a circular shift.
%
For example, for $n = 3$, there are $2^3 = 8$ binary strings,
but only four necklaces: 000, 001, 011 and 111,
since 010 and 100 are equivalent to 001,
so are 110 and 101 to 011.
%
The mapping is constructed as follows.
%
Given a cyclic polynomial, if the generator contains $x_k$,
  the $k$th character from the right of the necklace is 1,
  otherwise 0,
e.g. the above four necklaces correspond to
  $C_0$, $C_1$, $C_{12}$ and $C_{123}$, respectively.
%For example, if $n = 3$,
% the square-free polynomials for $000$, $100$, $110$ and $111$ are
%  $C_0 = 1$ (generator: 1),
%  $C_1 = x_1 + x_2 + x_3$ (generator: $x_1$),
%  $C_{12} = x_1 x_2 + x_2 x_3 + x_3 x_1$ (generator: $x_1 x_2$)
%  and $C_{123} = x_1 x_2 x_3$ (generator: $x_1 x_2 x_3$), respectively.
Thus,
%
%\begin{theorem}
the number $\NB(n)$ of the square-free cyclic polynomials $C_p$,
  %$(\vx)$ formed by the cycle points $\vx = \{x_1, \ldots, x_n\}$,
  is given by
\begin{equation}
  %\deg_\lambda A_n\lam =
  \NB(n) = (1/n) \textstyle\sum_{d|n} \varphi(n/d) 2^{\,d},
\label{eq:necklace}
\end{equation}
%
%\label{thm:necklace}
%\end{theorem}
%
where $\varphi(m)$ is Euler's totient function,
which gives the number of integers from 1 to $m$
  that are coprime to $m$,
  e.g.,
  $\varphi(1) = 1$,
  $\varphi(3) = 2$ for 1 and 2,
  $\varphi(6) = 2$ for 1 and 5.
By \refeq{xr},
  $\NB$ is also $\deg_\lambda A_n$.
The first few $\NB(n)$ are
2, 3, 4, 6, 8, 14, 20, 36, 60, 108, \dots, % 188, 352, 632, \ldots,
  from $n = 1$.

@@

%\begin{theorem}
Next the polynomial $P_n\lam$ can be written as
\begin{equation}
  P_n\lam = \textstyle\prod_l [\lambda - \Lambda(\vxn{n}^{(l)})],
\end{equation}
where $\vxn{n}^{(l)}$ denotes the $l$th $n$-cycle.
%
Thus, $\deg_\lambda P_n\lam$
  is equal to the number $\LB$ of the $n$-cycles,
  which is given by
  \cite{hao, hao2, lutzky}.
\begin{equation}
  %\deg_\lambda P_n\lam =
  \LB(n) = (1/n)\textstyle\sum_{d | n} \mu(n/d) 2^{\,d}.
\label{eq:lyndon}
\end{equation}
%\label{thm:lyndon}
%\end{theorem}
%
%
%\end{proof}
%
The first few $\LB$ are
2, 1, 2, 3, 6, 9, 18, 30, 56, 99, \dots %186, 335, 630, \ldots,
from $n = 1$.

Incidentally, $\LB$ is also the number of aperiodic necklaces of length $n$
(e.g., for $n = 4$,
  ``$0011$'' is aperiodic, but ``$0101$'' is not due to the shorter period 2).
%
Since an $n$-necklace is either aperiodic itself or
  a repeated aperiodic $d$-necklace (with $d|n$),
  we have
\begin{equation}
  \NB(n) = \textstyle\sum_{d|n} \LB(d).
\label{eq:necklacelyndon}
\end{equation}
%
Since the left side is the degree in $\lambda$ of $A_n\lam$,
  which accommodates all fixed points of $f^n$,
  %\refeq{necklacelyndon} shows that
  each distinct $d$-cycle (with $d|n$)
  contributes one and only one zero to $A_n\lam$.
%
%
We may thus write
%\begin{theorem}
  \begin{equation}
    A_n\lam = \textstyle\prod_{d|n} Q_d^{n/d}\lam,
    \label{eq:prod}
  \end{equation}
where $Q_d^{n/d}(\lambda) \equiv \prod_l
  \big[\lambda - \Lambda\big(\vxn{n}^{(l)}\big) \big]$
  with $\vxn{n}^{(l)}$ being the $n$ points
  corresponding to the $l$th $d$-cycle.
%
In order words, $\vxn{n}^{(l)}$ is $\vxn{d}^{(l)}$
  repeated for $c = n/d$ times:
\begin{equation}
  \vxn{n}^{(l)} =
  \{
    x_1^{(l)}, \ldots, x_d^{(l)},
    \; \ldots, \;
    x_1^{(l)}, \ldots, x_d^{(l)}
  \}
  = \vxn{d}^{(l)} * c.
\notag
\end{equation}
%
It follows that
$\Lambda(\vxn{n})$ = $[\Lambda(\vxn{d})]^c$.
By factorizing $\lambda - \Lambda^c$
  in terms of complex roots of unity, we get
\begin{align}
Q_d^{c}\lam
%  &= \prod_l
%    \Big\{
%      \lambda - \big[\Lambda\big(\vxn{d}^{(l)}\big) \big]^{c}
%    \Big\}
%    \notag \\
  &=
    \prod_l
    \Big\{
      \prod_{k=1}^c \,
      \big[
          e^{2k\pi i/c} \lambda^{1/c}
          - \Lambda\big(\vxn{d}^{(l)}\big)
      \big]
    \Big\},
    \notag
\end{align}
%
Now changing the order of the two products yields
\begin{equation}
Q_d^{c}\lam
  =  \prod_{k=1}^c P_d\big(e^{2k\pi i/c} \lambda^{1/c}\big),
\label{eq:QP}
\end{equation}
%
%
Despite the presence of $\lambda^{1/c}$,
  $Q_d^c\lam$ is a polynomial of $\lambda$,
  thanks to the invariance
  $\lambda \rightarrow e^{2\pi i} \lambda$;
and
  $\deg_\lambda Q_d^c\lam
    = \deg_\lambda P_d\lam = \LB(d)$.

%Therefore, the product $\prod_{cd=n} Q_d^c\lam$
%  can differ from $A_n\lam$ only by
%  a multiple.
%Since $Q_n^1\lam = P_n\lam$,
%  and the coefficient of highest power of $\lambda$ is always
%  unity in $A_n\lam$ \big[see the definition \refeq{xr}\big],
%  we know by induction that the coefficients of the highest power of $\lambda$
%  in all $P_n\lam$ and $Q_d^c\lam$
%  are also unities.
%So the multiple is one, hence \refeq{prod}.

\refeq{primfac} can now be derived from
  \refeqs{prod} and (\ref{eq:QP}),
%\begin{small}
\begin{align*}
 B_n^m (\lambda)
% & = \prod_{r = 1}^{m} A_n(e^{2l\pi i/m} \lambda^{1/m}) \\
 & = \prod_{k' = 1}^m \prod_{c | n} \prod_{k=1}^c
      P_{n/c}\big(
        e^{2k\pi i/c + 2 k' \pi i/(m c)}
        \lambda^{1/(m c)}
      \big) \\
 & = \prod_{c | n} \prod_{k''=1}^{m c}
    P_{n/c}(e^{2k''\pi i/(m c)} \lambda^{1/(m c)}) \\
 &
  = \prod_{c | n} Q_{n/c}^{m c}(\lambda).
\end{align*}
%\end{small}
The M\"obious inversion yields
\begin{equation}
Q_n^m\lam
= \textstyle\prod_{d|n} \big[
    B_d^{m n/d}\lam
  \big]^{\mu(n/d)}.
\notag
%\tag{$\ref{eq:primfac}'$}
\end{equation}
\refeq{primfac} is the special case of $m = 1$,
for $Q_n^1 = P_n$.

@@

Since the $n$-odd-cycles satisfy a polynomial of lower degree,
  the $n$-cycle polynomial $P_{2n}(\lambda)$
  (omitting the argument $r$)
  can always be factorized.
%
Suppose $\Lambda^o(\vx) \equiv \prod_{k = 1}^{n} f'(x_k)$ in the odd-cycle
 satisfies $P_{n}^o(\lambda^o) = 0$
\big[where $\lambda^o$ is the value of $\Lambda^o(\vx)$,
and $\lambda^o = \pm\sqrt \lambda$\,\big],
then $P^o_{n}(\sqrt \lambda) P^o_{n/2}(-\sqrt \lambda)$
is a factor of $P_{n}(\lambda)$.
%
For example, by solving the $n = 2$ odd-cycle,
  we have $- x_1 = r x_1 - {x_1}^3$, or ${x_1}^2 = r + 1$.
Since $\lambda^o = r - 3 {x_1}^2$,
$P^o_1(\lambda^o) = \lambda^o + 2 r + 3$.
%
The 2-cycle polynomial
  $P_2(\lambda) = -\big[(2r+3)^2 - \lambda\big]\,(\lambda + 2r^2-9)$,
  indeed contains
  $P^o_1(\sqrt \lambda) \, P^o_1(r, -\sqrt \lambda)
   = (\sqrt \lambda + 2 r + 3)(-\sqrt \lambda + 2 r + 3)$
  as a factor.
